---
title: "Descriptive Analysis"
author: ""
date: "2022-10-08"
output: html_document
---

TODO: Explanation of this part

```{r, setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
requiredPackages <- c("ggplot2","rapportools","plyr","cowplot", "ggpubr", "corrplot", "corrplot", "mlbench", "caret", "FactoMineR")
packages.check <- lapply(requiredPackages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

PROJECT_PATH="~/Documents/UPC/1Q/MultivariateAnalysis_MVVA/Labs/MultivariantAnalysis_IBMDataset"

setwd(PROJECT_PATH)

```

# Content Table

1.  Introduction

    1.1 Load Data

    1.2 DataFrame Basic Information

2.  Univariate analysis

    2.1 Categorical

    2.2 Numerical

3.  Bivariate analysis

    3.1 Contingency Table

4.  Outliers Detection

    4.1 Uni-variant

    4.2 Multi-variant

5.  Derivation of New Variables

6.  Exploratory Data Analysis

7.  Unbalanced Dataset

8.  Normalization

# 1. Introduction

## 1.1 Load Pre-processed data and useful functions

```{r, echo=TRUE}

load("./Data/df_preprocessed_IBM_Employees.Rdata") 


plotFreq <- function(df, feature, xlabel) {
  plot<-ggplot(df, aes(x=feature)) +
        geom_bar(aes(y =(..count..)/sum(..count..))) +
        geom_text(aes( label = scales::percent((..count..)/sum(..count..)),
                   y= ((..count..)/sum(..count..)) ), stat= "count", vjust = 1.25, color="#ffffff", size=3) +
        labs(x=xlabel, y='Percent')+
        scale_y_continuous(labels = scales::percent)
  return(plot)
}

plotPie <- function(feature, varName) {
  nb <- count(feature)[2];
  per <- c(round(nb/sum(nb)*100,2))$freq;
  names <- levels(feature);
  labs <- paste(names, per, "%");
  #jpeg(paste(plot_path, data_names[i], ".jpeg", sep = ""), height=800, width=1000)
  plot<-pie(summary(feature), labels = labs, main=paste("Pie Chart of ",varName));
  return(plot)
}

```

## 1.2 Dataframe basic information

Show a summary of the current dataframe and split by numeric and categorical variables.

```{r, echo=TRUE}

str(df)
summary(df)


numericalDf <- df[!sapply(df, is.factor)]
categoricalDf <- df[sapply(df, is.factor)]

```

# 2. Univariate analysis

## 2.1 Categorical variables

```{r, echo=False}

# we make a pie-chart for categorical variable with no order and binary
plotPie(df$Attrition, 'Attrition')
plotPie(df$Department, 'Department')
plotPie(df$EducationField, 'EducationField')
plotPie(df$Gender, 'Gender')
plotPie(df$JobRole, 'JobRole')
plotPie(df$MaritalStatus, 'MaritalStatus')
plotPie(df$OverTime, 'OverTime')

# we make an histogram for categorical variable with order
plotFreq(df, df$PerformanceRating, 'PerformanceRating')
plotFreq(df, df$BusinessTravel, 'BusinessTravel')
plotFreq(df, df$Education, 'Education')
plotFreq(df, df$EnvironmentSatisfaction, 'EnvironmentSatisfaction')
plotFreq(df, df$JobInvolvement, 'JobInvolvement')
plotFreq(df, df$JobLevel, 'JobLevel')
plotFreq(df, df$JobSatisfaction, 'JobSatisfaction')
plotFreq(df, df$RelationshipSatisfaction, 'RelationshipSatisfaction')
plotFreq(df, df$StockOptionLevel, 'StockOptionLevel')
plotFreq(df, df$WorkLifeBalance, 'WorkLifeBalance')

```

## 2.2 Numerical Variables

```{r, echo=False}

nonFactorDf <- df[!sapply(df, is.factor)]

i = 1
data_names <- attributes(nonFactorDf)$names
for (variable in nonFactorDf) {
  hist(variable, main = data_names[i])
  i = i+1
}

```

```{r, echo=False}

quantile(df$SalaryHikePerYear, prob=c(.1,.2,.3,.4,.5,.6,.7,.8,.9))
quantile(df$AverageYearsCompany, prob=c(.1,.2,.3,.4,.5,.6,.7,.8,.9))

```

# 3. Bivariate analysis

```{r, echo=False}

nonFactorDf <- df[!sapply(df, is.factor)]

#RateVariables 

pairs(df[,c("DailyRate", "HourlyRate", "MonthlyRate")])
pairs(df[,c("DailyRate", "HourlyRate", "MonthlyRate", "YearsAtCompany","MonthlyIncome")])
# podemos ver que la relacion entre ellas es nula

ggplot(df, aes(x = MonthlyRate, y = MonthlyIncome)) +
  geom_point()

#df$DailyRate / df$HourlyRate

ggplot(df, aes(x = 1:nrow(df), y = DailyRate / HourlyRate)) +
  geom_point()+
  labs(x = "Table Index")

```

## 3.1 Correlation study

```{r}

i=1
for (var1 in nonFactorDf[-length(nonFactorDf)]) {
  nonFactorDf3 <- nonFactorDf[-c(1:i)]
  j=1
  for(var2 in nonFactorDf3) {
    correlation <- cor(var1, var2)
    if(correlation > 0.75) {
      print(paste0(attributes(nonFactorDf[i])$names, " | ", attributes(nonFactorDf3[j])$names, " | correlation ", correlation))
    }
    j=j+1
  }
  i=i+1
}

#Highly corralated variables:
# - Total Working years - monthly income
# - Years with current manager - Years at the company - Years current role
# We should think of eliminating the variables that are highly correlated.
# Probably doing and stadistic test would be a good idea to know if we have to
# delete them or no.

```

## 3.2 Dependancy between Numerical variable and Response variable

```{r}
#Now we are going to do  ANOVA test wish is a is a type of statistical test used to determine if there is a statistically significant difference between two or more categorical groups by testing for differences of means using variance.
#We will do ANOVA test for combination of each numerical variable and the categorical response variable (Attrition).
#We will keep the numerical variable when the p-value of the ANOVA test is < 0.05, the cases where we are pretty sure that there is a significant difference of means (means of the numerical variable tested) because of the attrition .

i=1
for (var1 in nonFactorDf) {
  pvalue = oneway.test(var1~df$Attrition)$p.value
  if(pvalue < 0.05) {
    print(paste0(attributes(nonFactorDf[i])$names, " | Attrition | p-value : ", pvalue))
  }
  i = i+1
}

#ggplot(df, aes(x=Age, y = ..density.., color=Attrition)) +
#     geom_histogram()

ggplot(df, aes(x=Age, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=DailyRate, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=DistanceFromHome, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=TotalWorkingYears, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=TrainingTimesLastYear, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=YearsAtCompany, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=YearsInCurrentRole, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=YearsSinceLastPromotion, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=YearsWithCurrManager, color=Attrition)) +
     geom_density()

ggplot(df, aes(x=MonthlyIncome, color=Attrition)) +
     geom_density()

```

-So we can say that if an employee want to leave the company, he has probably a higher value than the mean for the following variable : DistanceFromHome. and a smaller value for : Age, DailyRate, TotalWorkingYears, TrainingTimesLastYear, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager. - We can expect that a Young Employee with little experience in the company has more risk of leaving the company.

## 3.3 Contingency table

```{r, echo=False, warning=FALSE}

# for all combination between two categorical variable, we make a chisq.test (independance test for categorical var) and we keep variable than are dependant (p-value of the chisq test < 0.05 => we are sure than there are dependant with > 95%)

i=1
for (var1 in factorDf[-length(factorDf)]) {
  factorDf3 <- factorDf[-c(1:i)]
  j=1
  for(var2 in factorDf3) {
    chisq_test <- chisq.test(var1, var2)
    if(chisq_test$p.value < 0.05) {
      print(paste0(attributes(factorDf[i])$names, " | ", attributes(factorDf3[j])$names, " | p-value ", chisq_test$p.value))
      print(table(var1,var2))
    }
    j=j+1
  }
  i=i+1
}

ggplot(data = factorDf) +
  geom_bar(aes(x = factor(OverTime), fill = factor(JobSatisfaction)), position = "fill")

ggplot(data = factorDf) +
  geom_bar(aes(x = factor(JobRole), fill = factor(Gender)), position = "fill")

ggplot(data = factorDf) +
  geom_bar(aes(x = factor(JobLevel), fill = factor(StockOptionLevel)), position = "fill")

ggplot(data = factorDf) +
  geom_bar(aes(x = factor(MaritalStatus), fill = factor(StockOptionLevel)), position = "fill")
```

-   We can see that the more an employee make over time the more he likes his job.
-   Women have a little bit more positions of responsibility (in proportion to their number)
-   Single employees have no StockOptionLevel. -Divorced employees have more StockOptionLevel than others employees
-   Despite a p-value \< 0.05, looking at the contingency tables we can say that they are not 100% dependent and therefore we cannot eliminate them.
-   We may group the variable Department and JobRole for create a new variable (categorical variable with the following levels : Representatives R&D, Representatives Sales, Healthcare Representative, Laboratory Technician, Manager Human Resources, Manager R&D, Manager Sales, Directors R&D, Research Scientist, Sales Executive)but we haven't do that because we prefer keep the information about the both variable rather than simplify the dataset by reduce his number of variable.
-   Sellers are more likely to want to quit the company
-   The following variables will be interesting to predict our response variable (Attrition) : Department, EducationField, EnvironmentSatisfaction, JobInvolvement, JobLevel, JobRole, JobSatisfaction, MaritalStatus, OverTime, StockOptionLevel, WorkLifeBalance.

## 3.4 Extra exploratory Data Analysis

```{r, echo=False}

# Correlation Matrix of numerical variables

corrplot(cor(nonFactorDf), method="number")
corrplot(cor(nonFactorDf), method="color")


corr <- round(cor(nonFactorDf), 1)

pairs(df[, c("MonthlyIncome", "YearsSinceLastPromotion",
             "TotalWorkingYears", "YearsAtCompany")])


# Comparation with attrition

hist_PercentSalaryHike <-ggplot(df, aes(x=PercentSalaryHike, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_TotalWorkingYears <-ggplot(df, aes(x=TotalWorkingYears, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_TrainingTimesLastYear <-ggplot(df, aes(x=TrainingTimesLastYear, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_YearsAtCompany <-ggplot(df, aes(x=YearsAtCompany, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_YearsInCurrentRole <-ggplot(df, aes(x=YearsInCurrentRole, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_YearsSinceLastPromotion <- ggplot(df, aes(x=YearsSinceLastPromotion, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_YearsWithCurrManager <- ggplot(df, aes(x=YearsWithCurrManager, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 1)

hist_MonthlyIncome <- ggplot(df, aes(x=MonthlyIncome, fill=Attrition, color=Attrition)) +
  geom_histogram(position="identity", alpha=0.5, binwidth = 350)


figure <- ggarrange(hist_PercentSalaryHike, hist_TotalWorkingYears, hist_TrainingTimesLastYear,
                    hist_YearsAtCompany, hist_YearsInCurrentRole, hist_YearsSinceLastPromotion, 
                    hist_YearsWithCurrManager, hist_MonthlyIncome,
                    ncol = 2, nrow = 4)
figure



# Scatterplot

ggplot(df, aes(x=MonthlyIncome, y=YearsAtCompany, color=Attrition)) + 
  geom_point() +
  geom_smooth()


ggplot(df, aes(x=YearsSinceLastPromotion, y=YearsAtCompany, color=Attrition)) + 
  geom_point() +
  geom_smooth()

ggplot(df, aes(x=YearsInCurrentRole, y=MonthlyIncome, color=Attrition)) + 
  geom_point()+
  geom_smooth()



# EDU POTSER REPETITS
## PERCENTSALARYHIKE CORRELATED WITH PERFORMANCERATING!!

ggplot(df, aes(x=Age, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=DailyRate, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=DistanceFromHome, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=MonthlyIncome, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=NumCompaniesWorked, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=PercentSalaryHike, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=TotalWorkingYears, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=YearsAtCompany, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=YearsInCurrentRole, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=YearsSinceLastPromotion, y=Attrition)) +
  geom_boxplot()

ggplot(df, aes(x=YearsWithCurrManager, y=Attrition)) +
  geom_boxplot()


# OTHERS

ggplot(df, aes(x=MonthlyIncome, y=JobLevel, fill=Gender)) +
  geom_boxplot()

ggplot(df, aes(x=PercentSalaryHike, y=PerformanceRating, fill=Gender)) +
  geom_boxplot()

ggplot(df, aes(x=MonthlyIncome, y=EducationField, fill=Gender)) +
  geom_boxplot()

ggplot(df, aes(x=MonthlyIncome, y=EducationField, fill=MaritalStatus)) +
  geom_boxplot()

ggplot(df, aes(x=MonthlyIncome, y=Attrition, fill=Gender)) +
  geom_boxplot()


ggplot(df, aes(StockOptionLevel, ..count..)) + geom_bar(aes(fill = MaritalStatus), position = "dodge")

ggplot(df, aes(StockOptionLevel, ..count..)) + geom_bar(aes(fill = Attrition), position = "dodge")

ggplot(df, aes(JobRole, ..count..)) + geom_bar(aes(fill = Gender), position = "dodge")

summary(df$Attrition)
summary(df$Attrition[df$MonthlyIncome<4000])

```
