---
title: "MCA"
output:
  word_document: default
  html_document: default
date: "2022-10-16"
---


```{r, setup, include=FALSE, echo=TRUE}
#knitr::opts_chunk$set(echo = TRUE)
library(ggcorrplot)
library(ggplot2)
library(dplyr)
library(dlookr)
library(Factoshiny)
library(FactoMineR)
library(factoextra)
```

## 1.1 Load Data

```{r, echo=TRUE}
df <- read.csv("C:/Users/David/Downloads/WA_Fn-UseC_-HR-Employee-Attrition.csv")

#identical(Attrition, HELLOS)

# no va todavia CleanAttrition <- read.csv("C:/Users/David/Desktop/Máster/MVA/Project/CleanAttrition.csv")
```
# 2. Formatting issues

-   Check that Categorical variables are correctly processed by R, so all interpreted as factors.
-   Set short-meaningful level names for the categorical variables.
-   Reduce number of levels if has sense.

```{r, echo=False}
# Remove useless attributes
df <- subset(df, select = -c(EmployeeCount, StandardHours,Over18, EmployeeNumber))
#Transform to factor
df$Attrition=as.factor(df$Attrition)

df$BusinessTravel=as.factor(df$BusinessTravel)

df$Department=as.factor(df$Department)

df$Education=as.factor(df$Education)
levels(df$Education) <- c('Below College', 'College', 'Bachelor', 'Master', 'Doctor')

df$EducationField=as.factor(df$EducationField)

df$EnvironmentSatisfaction=as.factor(df$EnvironmentSatisfaction)
levels(df$EnvironmentSatisfaction) <- c('Low', 'Medium', 'High', 'Very High')

df$Gender=as.factor(df$Gender)

df$JobInvolvement=as.factor(df$JobInvolvement)
levels(df$JobInvolvement) <- c('Low', 'Medium', 'High', 'Very High')

df$JobLevel=as.factor(df$JobLevel)
levels(df$JobLevel) <- c('Intern', 'Junior', 'Senior', 'Manager', 'Director')

# Reduce number of levels from 9 to 7
df$JobRole=as.factor(df$JobRole)
levels(df$JobRole) <- c("Representatives", "H.R.", "Lab. Tech.", "Manager", "Directors", "Directors", "Research Sci.", "Sales Executive", "Representatives")

df$JobSatisfaction=as.factor(df$JobSatisfaction)
levels(df$JobSatisfaction) <- c('Low', 'Medium', 'High', 'Very High')

df$MaritalStatus=as.factor(df$MaritalStatus)

df$OverTime=as.factor(df$OverTime)

df$PerformanceRating=factor(df$PerformanceRating, levels = c(1, 2, 3, 4))
levels(df$PerformanceRating) <- c('Low', 'Good', 'Excellent', 'Outstanding')

df$RelationshipSatisfaction=as.factor(df$RelationshipSatisfaction)
levels(df$RelationshipSatisfaction) <- c('Low', 'Medium', 'High', 'Very High')

df$StockOptionLevel=as.factor(df$StockOptionLevel)
levels(df$StockOptionLevel) <- c('No', 'Few', 'Medium', 'High')

df$WorkLifeBalance=as.factor(df$WorkLifeBalance)
levels(df$WorkLifeBalance) <- c('Bad', 'Good', 'Better', 'Best')

Attrition <- df#clean dataset
#colnames(Attrition)

```
### 3 MCA with all variables
Active individuals (rows 1:1470): Individuals that are used in the multiple correspondence analysis.
Active variables (all categorical data) : Variables that are used in the MCA.
Supplementary variables: They don’t participate to the MCA. The coordinates of these variables will be predicted.
Supplementary quantitative variables (quanti.sup): all 
Supplementary qualitative variables (quali.sup:none
The null hypothesis of the Chi-Square test is that no relationship exists on the categorical variables in the population; they are independent.
```{r, echo=False}
#Split cat and num variables
cat_variables <- sapply(Attrition, is.factor)
num_variables <- sapply(Attrition, is.numeric)
cat.df <- Attrition[, cat_variables]
num.df <- Attrition[, num_variables]
#num of variables p = 17 and num of categories/ modalities k = 65 max inertia per dimension 35%, the mean intertia is 1/p = 0.0588 till dim 22
# Apply MCA function
res.mca <-MCA(cat.df, ncp = 8, graph = FALSE)


```
### 3.1 Visualization and Interpretation
We can see from the plots above that for both dimensions, the variables with relevant contribution are JobLevel, EducationField, Department and JobRole. Their influnce is higher for the first dimension. There seems to be a direct relation with d1 for the categories related to Human resources and also but not as much with the research ones. There is also and inverse relation with the sales related categories for D1.
```{r}
#get_eigenvalue(res.mca): Extract the eigenvalues/variances retained by each dimension (axis)
eigen.val <- get_eigenvalue(res.mca)
eigen.val
#fviz_eig(res.mca): Visualize the eigenvalues/variances
fviz_eig(res.mca, addlabels = TRUE,main = "MCA-Scree plot", ylim =c(0,6), ncp = 15)
#fviz_mca_biplot(res.mca): Make a biplot of rows and columns.
fviz_mca_biplot(res.mca,ggtheme = theme_minimal())
#get_mca_ind(res.mca), get_mca_var(res.mca): Extract the results for individuals and variables, respectively.
var <-get_mca_var(res.mca)
# Coordinates
head(var$coord)
# Cos2: quality on the factore map
head(var$cos2)
# Contributions to the principal components
head(var$contrib)
#fviz_mca_ind(res.mca), fviz_mca_var(res.mca): Visualize the results for individuals and variables, respectively.
fviz_mca_var(res.mca, choice = "mca.cor", 
            repel = TRUE, # Avoid text overlapping (slow)
            ggtheme = theme_minimal())
#this for only variable categories
fviz_mca_var(res.mca, axes = c(3,4), choice = "var.cat", 
             repel = TRUE, select.var = list(  contrib = 10),
            ggtheme = theme_minimal())
```



```{r}
# Cos2 of variable categories on Dim.1 and Dim.2
fviz_cos2(res.mca, choice = "var", axes = 1:2, top = 10)
# Color by contrin values: quality on the factor map
fviz_mca_var(res.mca, col.var = "contrib",   
             select.var = list(contrib = 15),
              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())
# Contributions of rows to dimension 1
fviz_contrib(res.mca, choice = "var", axes = 1, top = 15)
# Contributions of rows to dimension 2
fviz_contrib(res.mca, choice = "var", axes = 2, top = 15)
# Contributions of rows to dimension 3
fviz_contrib(res.mca, choice = "var", axes = 3, top = 15)
# Contributions of rows to dimension 4
fviz_contrib(res.mca, choice = "var", axes = 4, top = 15)
fviz_mca_var(res.mca, axes = c(3,4), col.var = "contrib",   
             select.var = list(contrib = 15),
              gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             
             
             repel = TRUE, # Avoid text overlapping
             ggtheme = theme_minimal())
#Dimension description
res.desc <- dimdesc(res.mca, axes = c(1,2,3,4,5,6,7,8))
# Description of dimension 1res.desc[[1]]

# Description of dimension 2res.desc[[2]]

# Description of dimension 3res.desc[[3]]

# Description of dimension 4res.desc[[4]]

res.desc$`Dim 1`$quali
res.desc$`Dim 2`$quali
res.desc$`Dim 3`$quali
res.desc$`Dim 4`$quali

```

```{r}
fviz_mca_ind(res.mca, label="none",axes = c(3,4), habillage=1,addEllipses=TRUE, ellipse.level=0.95)#Attritio
fviz_mca_ind(res.MCA, label="none",axes = c(1,2), habillage=5,addEllipses=TRUE, ellipse.level=0.95)#dep
fviz_mca_ind(res.MCA, label="none",axes = c(1,2), habillage=13,addEllipses=TRUE, ellipse.level=0.95)#JL
fviz_mca_ind(res.MCA, label="none",axes = c(1,2), habillage=14,addEllipses=TRUE, ellipse.level=0.95)#JR
fviz_mca_ind(res.MCA, label="none",axes = c(2,4), habillage=13,addEllipses=TRUE, ellipse.level=0.95)
fviz_mca_ind(res.MCA, label="none",axes = c(3,4), habillage=24,addEllipses=TRUE, ellipse.level=0.95)#stock
fviz_mca_ind(res.MCA, label="none",axes = c(8,3), habillage=2,addEllipses=TRUE, ellipse.level=0.95)
fviz_mca_ind(res.MCA, label="none",axes = c(3,5), habillage=2,addEllipses=TRUE, ellipse.level=0.95)
fviz_mca_ind(res.MCA, label="none",axes = c(3,4), habillage=16,addEllipses=TRUE, ellipse.level=0.95)#marital
fviz_mca_ind(res.MCA2, label="none",axes = c(6,2), habillage=27,addEllipses=TRUE, ellipse.level=0.95)#balance
fviz_mca_ind(res.MCA2, label="none",axes = c(3,2), habillage=12,addEllipses=TRUE, ellipse.level=0.95)#involvement
fviz_mca_ind(res.MCA2, label="none",axes = c(1,2), habillage=20,addEllipses=TRUE, ellipse.level=0.95)#overtime  
fviz_mca_ind(res.MCA2, label="none",axes = c(1,2), habillage=3,addEllipses=TRUE, ellipse.level=0.95)#travel 

```
### 4 Specific MCA
Active individuals (rows 1:1470): Individuals that are used in the multiple correspondence analysis.
Active variables (all categories) : Variables that are used in the MCA.
Supplementary variables: They don’t participate to the MCA. The coordinates of these variables will be predicted.
Supplementary quantitative variables (quanti.sup): All quantitative variables.
Supplementary qualitative variables (quali.sup: Columns 3 and 4 corresponding to the columns Sick and Sex, respectively. This factor variables will be used to color individuals by groups.
The null hypothesis of the Chi-Square test is that no relationship exists on the categorical variables in the population; they are independent.
```{r}
# DEEP ANALYSIS # ATTRITION SEPARADO YES NO CON CATEGORIAS CON SENTIDO
#division por diferentes departamentos d1 y d2 en este caso estan relacionados con esto 
res.MCA<-MCA(Attrition,quanti.sup=c(1,4,6,11,17,18,19,21,25,26,28,29,30,31),graph=FALSE, ncp = 8)
plot.MCA(res.MCA,selectMod= 'contrib  17 ',habillage=5,label =c('var'))#DEP
plot.MCA(res.MCA,selectMod= 'contrib  17 ',habillage=8,label =c('var'))#EDUCATION FIELD
plot.MCA(res.MCA,selectMod= 'contrib  17 ',habillage=14,label =c('var'))#JOB ROLE
#DIM 2 4 JOB LEVEL
plotellipses(res.MCA,keepvar=13,axes=c(2,4),label ='quali')
#DIM 3 Y 2 DEPARTMENS SE VE BONITO
plotellipses(res.MCA,keepvar=5,axes=c(3,2),label ='quali')
# DIMENSIONES 3 Y 4 RELACIONADAS CON 
plot.MCA(res.MCA,axes=c(3,4),selectMod= 'contrib  17 ',habillage=24,label =c('var'))#STOCK
plot.MCA(res.MCA,axes=c(3,4),selectMod= 'contrib  17 ',habillage=16,label =c('var'))#MARITAL
plot.MCA(res.MCA,axes=c(3,4),selectMod= 'contrib  17 ',habillage=14,label =c('var'))#JOB ROLE
plot.MCA(res.MCA,axes=c(3,4),selectMod= 'contrib  17 ',habillage=13,label =c('var'))#JOB LVL CORRELATED WITH PREVIO
plot.MCA(res.MCA,axes=c(3,4),selectMod= 'contrib  17 ',habillage=2,label =c('var'))# ATTRITION!!
#DIMENSIONES 5 Y 7
plot.MCA(res.MCA,axes=c(5,7),selectMod= 'contrib  11 ',habillage=20,label ='none')#OVERTIME
#DIMENSIONES 6 Y 8 GENDER
plot.MCA(res.MCA,axes=c(6,8),selectMod= 'contrib  11 ',habillage=10,label =c('var')) # POCA INERTIA
#DIMENSITONES 8 Y 7 JOBINVO
plotellipses(res.MCA,keepvar=12,axes=c(7,8),label ='quali')
#STOCKOPTION 7 Y 3/4
plotellipses(res.MCA,keepvar=24,axes=c(3,7),label ='quali')
plot.MCA(res.MCA,axes=c(7,4),selectMod= 'contrib  11 ',habillage=24,label =c('var'))
plot.MCA(res.MCA,axes=c(7,3),selectMod= 'contrib  11 ',habillage=24,label =c('var'))

#ATTRITION PLOTS
plot.MCA(res.MCA,axes=c(8,3),selectMod= 'contrib  11 ',habillage=2,label =c('var'))
plot.MCA(res.MCA,axes=c(3,4),selectMod= 'contrib  17 ',habillage=2,label =c('var'))
plot.MCA(res.MCA,axes=c(7,3),selectMod= 'contrib  11 ',habillage=2,label =c('var'))
plot.MCA(res.MCA,axes=c(5,3),selectMod= 'contrib  11 ',habillage=2,label =c('var'))
#this presents overtime with job satisfaction with filtered MCA
res.MCA2<-MCA(Attrition, ncp = 8, quanti.sup=c(1,4,6,11,17,18,19,21,25,26,28,29,30,31),quali.sup=c(2,5,7,8,13,14,16,24),graph=FALSE)
plot.MCA(res.MCA2,selectMod= 'contrib  9 ',habillage=20,label =c('var','quali.sup'))#overtime job satis
plot.MCA(res.MCA2,selectMod= 'contrib  9 ',habillage=3,label =c('var','quali.sup')) #travel 
plot.MCA(res.MCA2,axes=c(6,2),selectMod= 'contrib  9 ',habillage=27,label =c('quali.sup'))# dim 2 and 6 worklife balan
plot.MCA(res.MCA2,axes=c(3,2),selectMod= 'contrib  9 ',habillage=12,label =c('quali.sup')) # dim 2 and 3 JobInv invers
plot.MCA(res.MCA2,axes=c(3,2),selectMod= 'contrib  9 ',habillage=15,label =c('quali.sup'))# dim 2 and 3 Jobsatisf 


#PROGRAMA
#AttritionMCA <-Factoshiny(Attrition)
#Factoshiny(AttritionMCA)
```





