---
title: "Missing values"
output: pdf_document
date: "2022-10-14"
---

```{r, setup, include=FALSE, echo=TRUE}
#knitr::opts_chunk$set(echo = TRUE)
library(mice)
library(ggcorrplot)
library(naniar)
library(ggplot2)
library(finalfit) 
library(visdat)
library(dplyr)
library(dlookr)
library(DMwR2) #knn
library(GGally)
library(VIM)
library(missForest)
library(FactoMineR)
library(shiny)
library(FactoInvestigate)
library(Factoshiny)
library(factoextra)
#mixtools, MVN, robustbase
```

# Content Table

1.  Introduction

    1.1 Load Data

    1.2 DataFrame Basic Information

2.  Formatting issues

3.  MFA

    3.1 Summary of outputs / Eigenvalues
    3.2 1st/2nd dimension - Attrition
    3.3 3rd/4th dimension - JobRole

# 1. Introduction

## 1.1 Load Data

```{r, echo=TRUE}
df <- read.csv("C:/Users/Tony Vitas/Downloads/WA_Fn-UseC_-HR-Employee-Attrition.csv")
```

## 1.2 Dataframe basic information

Show a summary of the current dataframe.

```{r, echo=TRUE}
explanatory = c("Age",
                "DailyRate",
                "DistanceFromHome",
                "HourlyRate",
                "MonthlyIncome",
                "MonthlyRate",
                "NumCompaniesWorked",
                "PercentSalaryHike",
                "TotalWorkingYears",
                "TrainingTimesLastYear",
                "YearsAtCompany",
                "YearsInCurrentRole",
                "YearsSinceLastPromotion",
                "YearsWithCurrManager",
                "Gender",
                "OverTime",
                "BusinessTravel",
                "Department",
                "Education",
                "EducationField",
                "EnvironmentSatisfaction",
                "JobInvolvement",
                "JobLevel",
                "PerformanceRating",
                "JobRole",
                "JobSatisfaction",
                "MaritalStatus",
                "RelationshipSatisfaction",
                "StockOptionLevel",
                "WorkLifeBalance")
dependent = "Attrition"
str(df)
head(df)
df%>% ff_glimpse (dependent, explanatory)# best
vis_dat(df)
```

# 2. Formatting issues

-   Check that Categorical variables are correctly processed by R, so all interpreted as factors.
-   Set short-meaningful level names for the categorical variables.
-   Reduce number of levels if has sense.

```{r, echo=False}
# Remove useless attributes
df <- subset(df, select = -c(EmployeeCount, StandardHours,Over18, EmployeeNumber))
#Transform to factor
df$JobInvolvement=as.factor(df$JobInvolvement)
df$JobLevel=as.factor(df$JobLevel)
df$PerformanceRating=as.factor(df$PerformanceRating)
df$JobRole = as.factor(df$JobRole)
levels(df$JobRole) <- c(1,2,3,4,5,6,7,8,9)
df$MaritalStatus=as.factor(df$MaritalStatus)
levels(df$MaritalStatus) <- c(1,2,3)
df$StockOptionLevel=as.factor(df$StockOptionLevel)
df$WorkLifeBalance=as.factor(df$WorkLifeBalance)
df$RelationshipSatisfaction=as.factor(df$RelationshipSatisfaction)
df$JobSatisfaction=as.factor(df$JobSatisfaction)
#Convert department to factors
df$Department=as.factor(df$Department)
levels(df$Department) <- c(1, 2, 3)
#Convert education to factors
df$Education=as.factor(df$Education)
#Convert education field to factors
df$EducationField=as.factor(df$EducationField)
levels(df$EducationField) <- c(1, 2, 3, 4, 5, 6)
#Convert environment satisfaction to factors
df$EnvironmentSatisfaction=as.factor(df$EnvironmentSatisfaction)
#Convert job satisfaction to factors
df$JobSatisfaction=as.factor(df$JobSatisfaction)
# Convert Business Travel to factors
df$BusinessTravel= as.factor(df$BusinessTravel)
levels(df$BusinessTravel) <- c(1, 2, 3)
# Convert Female/Male to Binary (as factor)
df$Gender=as.factor(df$Gender)
levels(df$Gender)<- c(0,1)
# Convert Yes/No to Binary (as factor)
df$Attrition = as.factor(df$Attrition)
levels(df$Attrition)<- c(0,1)
df$OverTime=as.factor(df$OverTime)
levels(df$OverTime) <-c(0,1)
Attrition_NoRandNA <- df#clean dataset
```

# 3. MFA

# 3.1 Eigenvalues // Variances
newDF <- Attrition_NoRandNA[,c("Age","NumCompaniesWorked","TotalWorkingYears","DistanceFromHome","YearsInCurrentRole","YearsAtCompany","MonthlyIncome","PercentSalaryHike","Gender","MaritalStatus","Attrition","OverTime","StockOptionLevel","Department","JobRole","JobLevel","BusinessTravel","JobSatisfaction","EnvironmentSatisfaction","RelationshipSatisfaction","WorkLifeBalance")]
res.MFA<-MFA(newDF,group=c(1,2,3,2,2,1,1,1,4,4), type=c("s","s","s","s","n","n","n","n","n","n"),name.group=c("GD_num","BK_num","JW_num","JE_num","GD_cat","Attrition","JW_cat","JE_cat","JC_cat","PSS_cat"),num.group.sup=c(1,5,10),graph=FALSE)

MFA.eig <- get_eigenvalue(res.MFA)
head(MFA.eig)

fviz_screeplot(res.MFA)


# 3.2 Graph of variables
# 3.2.1 Groups of variables
MFA.group <- get_mfa_var(res.MFA, "group")

head(MFA.group$coord) #Coordinate of groups
head(MFA.group$cos2) #Quality of representation on the factore map
head(MFA.group$contrib) #Contributions to the dimensions

fviz_mfa_var(res.MFA, "group") #Variables group plot

fviz_contrib(res.MFA, "group", axes=1) #Group contribution to Dim1

fviz_contrib(res.MFA, "group", axes=2) #Group contribution to Dim2

# 3.2.2 Quantitative variables

MFA.quanti <- get_mfa_var(res.MFA, "quanti.var")

head(MFA.quanti$coord) #Coordinate of groups
head(MFA.quanti$cos2) #Quality of representation on the factore map
head(MFA.quanti$contrib) #Contributions to the dimensions

fviz_mfa_var(res.MFA, "quanti.var", palette = "jco", color.var.sup = "violet", repel = TRUE, geom = c("point","text"), legend = "bottom") #Correlation circle for quanti var Dim1 and 2

fviz_mfa_var(res.MFA, "quanti.var", palette = "jco", color.var.sup = "violet", repel = TRUE, geom = c("point","text"), legend = "bottom", axes = c(3,4)) #Correlation circle for quanti var Dim3 and 4

fviz_contrib(res.MFA, choice = "quanti.var", axes=1, top = 10, palette = "jco") #Contribution to Dim1

fviz_contrib(res.MFA, choice = "quanti.var", axes=2, top = 10, palette = "jco") #Contribution to Dim2

fviz_contrib(res.MFA, choice = "quanti.var", axes=3, top = 10, palette = "jco") #Contribution to Dim3

fviz_contrib(res.MFA, choice = "quanti.var", axes=4, top = 10, palette = "jco") #Contribution to Dim4

fviz_mfa_var(res.MFA, "quanti.var", col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), geom = c("point", "text"), legend = "bottom")


# 3.2.3 Individuals

MFA.ind <- get_mfa_ind(res.MFA)

## 1st/2nd Dim
fviz_mfa_ind(res.MFA, habillage = "Attrition", palette = c("#00AFBB","#E7B800"), addEllipses = TRUE, repel = TRUE)

fviz_mfa_ind(res.MFA, habillage = "OverTime", palette = c("#00AFBB","#E7B800"), addEllipses = TRUE, repel = TRUE)

fviz_ellipses(res.MFA, c("Attrition","OverTime"), repel = TRUE, ellipse.type = "t")

## 3rd/4th Dim
plot.MFA(res.MFA, choix="ind",axes=c(3,4),lab.par=FALSE,select='contrib  500',habillage='JobRole',title="Individual factor map",cex=0.9,cex.main=0.9,cex.axis=0.9)


# 3.2.4 Partial Axes
plot.MFA(res.MFA, choix="axes",ncp=1,habillage='group',cex=1.05,cex.main=1.05,cex.axis=1.05)




