---
title: "Missing values"
output: pdf_document
date: "2022-10-14"
---

```{r, setup, include=FALSE, echo=TRUE}

requiredPackages <- c("mice", "ggcorrplot", "naniar", "ggplot2", "finalfit",
                      "visdat", "dplyr", "dlookr", "DMwR2", "GGally", "VIM", "missForest")
packages.check <- lapply(requiredPackages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
   }
})


```

# Content Table

1.  Introduction

    1.1 Load Data

    1.2 DataFrame Basic Information

2.  Formatting issues

3.  Check missing values

3.1 Generate not random NA MNAR

3.2 Generate random NA MCAR

4.  Missing values imputation

    4.1 Knn method for MNAR

    4.2 Knn method for MCAR

    4.3 Random Forest method for MNAR

    4.4 Random Forest method for MCAR

    4.5 mice method for MNAR

    4.6 mice method for MCAR

5.  Final Plots and Graphs

# 1. Introduction

## 1.1 Load Data

```{r, echo=TRUE}

PROJECT_PATH="~/Documents/UPC/1Q/MultivariateAnalysis_MVVA/Labs/MultivariantAnalysis_IBMDataset"

setwd(PROJECT_PATH)
df <- read.csv("./Data/IBM_Employees.csv")

```

## 1.2 Dataframe basic information

Show a summary of the current dataframe.

```{r, echo=TRUE}
explanatory = c("Age",
                "DailyRate",
                "DistanceFromHome",
                "HourlyRate",
                "MonthlyIncome",
                "MonthlyRate",
                "NumCompaniesWorked",
                "PercentSalaryHike",
                "TotalWorkingYears",
                "TrainingTimesLastYear",
                "YearsAtCompany",
                "YearsInCurrentRole",
                "YearsSinceLastPromotion",
                "YearsWithCurrManager",
                "Gender",
                "OverTime",
                "BusinessTravel",
                "Department",
                "Education",
                "EducationField",
                "EnvironmentSatisfaction",
                "JobInvolvement",
                "JobLevel",
                "PerformanceRating",
                "JobRole",
                "JobSatisfaction",
                "MaritalStatus",
                "RelationshipSatisfaction",
                "StockOptionLevel",
                "WorkLifeBalance")
dependent = "Attrition"
str(df)
head(df)
df%>% ff_glimpse (dependent, explanatory)# best
vis_dat(df)
```

# 2. Formatting issues

-   Check that Categorical variables are correctly processed by R, so all interpreted as factors.
-   Set short-meaningful level names for the categorical variables.
-   Reduce number of levels if has sense.

```{r, echo=False}
# Remove useless attributes
df <- subset(df, select = -c(EmployeeCount, StandardHours,Over18, EmployeeNumber))
#Transform to factor
df$JobInvolvement=as.factor(df$JobInvolvement)
df$JobLevel=as.factor(df$JobLevel)
df$PerformanceRating=as.factor(df$PerformanceRating)
df$JobRole = as.factor(df$JobRole)
levels(df$JobRole) <- c(1,2,3,4,5,6,7,8,9)
df$MaritalStatus=as.factor(df$MaritalStatus)
levels(df$MaritalStatus) <- c(1,2,3)
df$StockOptionLevel=as.factor(df$StockOptionLevel)
df$WorkLifeBalance=as.factor(df$WorkLifeBalance)
df$RelationshipSatisfaction=as.factor(df$RelationshipSatisfaction)
df$JobSatisfaction=as.factor(df$JobSatisfaction)
#Convert department to factors
df$Department=as.factor(df$Department)
levels(df$Department) <- c(1, 2, 3)
#Convert education to factors
df$Education=as.factor(df$Education)
#Convert education field to factors
df$EducationField=as.factor(df$EducationField)
levels(df$EducationField) <- c(1, 2, 3, 4, 5, 6)
#Convert environment satisfaction to factors
df$EnvironmentSatisfaction=as.factor(df$EnvironmentSatisfaction)
#Convert job satisfaction to factors
df$JobSatisfaction=as.factor(df$JobSatisfaction)
# Convert Business Travel to factors
df$BusinessTravel= as.factor(df$BusinessTravel)
levels(df$BusinessTravel) <- c(1, 2, 3)
# Convert Female/Male to Binary (as factor)
df$Gender=as.factor(df$Gender)
levels(df$Gender)<- c(0,1)
# Convert Yes/No to Binary (as factor)
df$Attrition = as.factor(df$Attrition)
levels(df$Attrition)<- c(0,1)
df$OverTime=as.factor(df$OverTime)
levels(df$OverTime) <-c(0,1)
Attrition_NoRandNA <- df#clean dataset
```

### 3 Check missing values

```{r, echo=False, warning=FALSE}
# Count missing values in colums
mis_col = colSums(is.na(df))
mis_col
# Count missing values in rows
mis_row = rowSums(is.na(df))
mis_row
#MORE Options used later
#countNA(df)
#vis_miss(df)
#gg_miss_upset(df) 
#pMiss <- function(x){sum(is.na(x))/length(x)*100}
#apply(df,2,pMiss)
#md.pattern(df)
#aggr_plot <- aggr(df, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE)

```

## 3.1 Generate not random NA MNAR

```{r, echo=False}

## Imput NA  NOT AT random MNAR

# Select some columns:
# 70% (66.66%) Numerical - 4 cols
# 1-  Age
# 13- HourlyRate
# 21- NumCompaniesWorked
# 32- YearsAtCompany

# 30% (33.33%) Categorical - 2 cols
# 12- Gender 0  female 
# 22- OverTime 0 yes 

# Select 2,6% miss data -> size= 200

individuals_idx = sample(nrow(df), size=200)
Attrition_NoRandNA[individuals_idx,c("Age", "HourlyRate","NumCompaniesWorked",
                      "YearsAtCompany", "Gender", "OverTime")] = NA
# NA's for each variable. Array of 1471 values
x <- colSums(is.na(df) | df == "")
sum(x)
x
# NA's for each row. Array of 1471 values
y <- rowSums(is.na(df) | df == "")
sum(y)
vis_miss(Attrition_NoRandNA)
gg_miss_upset(Attrition_NoRandNA) 
pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(Attrition_NoRandNA,2,pMiss)
md.pattern(Attrition_NoRandNA)
aggr_plot <- aggr(Attrition_NoRandNA, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE)
### Select the columns with NA
colsNA <- c("Age","HourlyRate", "NumCompaniesWorked", "YearsAtCompany","Gender","OverTime")
## Set a data frame with no NA
NONA <-na.omit(df[colsNA])
identical(NONA, df[colsNA]) #comprobar que son identicos
# Supuesto test de little
mcar_test(Attrition_NoRandNA) #so it says data is MCAR lol
```

### 3.2 Generate random NA MCAR

```{r, echo=False}
#Generate MCAR
Attrition_missRand <- prodNA(df, noNA = 0.026)# 2,6 % missing values
countNA(Attrition_missRand)
vis_miss(Attrition_missRand)
gg_miss_upset(Attrition_missRand)
vis_miss(Attrition_missRand)
pMiss <- function(x){sum(is.na(x))/length(x)*100}
apply(Attrition_missRand,2,pMiss)
md.pattern(Attrition_missRand)
aggr_plot <- aggr(Attrition_missRand, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE)

```

# 4. Missing values imputation

## 4.1 Knn method for MNAR

```{r, echo=False}
#DMwR2 library
df_imp_knn1NR = knnImputation(Attrition_NoRandNA[colsNA], k=1)
df_imp_knn2NR = knnImputation(Attrition_NoRandNA[colsNA], k=2)
df_imp_knn3NR = knnImputation(Attrition_NoRandNA[colsNA], k=3)
df_imp_knn4NR = knnImputation(Attrition_NoRandNA[colsNA], k=4)
df_imp_knn5NR = knnImputation(Attrition_NoRandNA[colsNA], k=5)
summary(df_imp_knn5NR)

#VIM library
Attrition_KNNNR <- kNN(Attrition_NoRandNA[,colsNA])
histMiss(Attrition_KNNNR, delimiter="imp", selection="any")
marginplot(Attrition_KNNNR[,c("Age","HourlyRate")], delimiter="imp",
           alpha = 0.8, 
           col = c("lightblue", "green", "black"))
#plots
plot(density(NONA$Age))
lines(density(df_imp_knn1NR$Age),col = 2 )
lines(density(df_imp_knn2NR$Age),col = 3 )
lines(density(Attrition_KNNNR$Age),col = 4 )
plot(density(NONA$NumCompaniesWorked))
lines(density(df_imp_knn4NR$NumCompaniesWorked),col = 5 )
lines(density(df_imp_knn5NR$NumCompaniesWorked),col = 6 )
lines(density(Attrition_KNNNR$NumCompaniesWorked),col = 4 )
```

## 4.2 Knn method for MCAR

```{r, echo=False}
#DMwR2 library
df_imp_knn1RandNA = knnImputation(Attrition_missRand[colsNA], k=1)
df_imp_knn2RandNA = knnImputation(Attrition_missRand[colsNA], k=2)
df_imp_knn3RandNA = knnImputation(Attrition_missRand[colsNA], k=3)
df_imp_knn4RandNA = knnImputation(Attrition_missRand[colsNA], k=4)
df_imp_knn5RandNA = knnImputation(Attrition_missRand[colsNA], k=5)
summary(df_imp_knn5RandNA)

#VIM library
Attrition_KNNRNA <- kNN(Attrition_missRand[,colsNA])
histMiss(Attrition_KNNRNA, delimiter="imp", selection="any")
marginplot(Attrition_KNNRNA[,c("Age","HourlyRate")], delimiter="imp",
           alpha = 0.8, 
           col = c("lightblue", "green", "black"))
#plots
plot(density(NONA$Age))
lines(density(df_imp_knn1RandNA$Age),col = 2 )
lines(density(df_imp_knn2RandNA$Age),col = 3 )
lines(density(Attrition_KNNRNA$Age),col = 4 )
plot(density(NONA$NumCompaniesWorked))
lines(density(df_imp_knn4RandNA$NumCompaniesWorked),col = 5 )
lines(density(df_imp_knn5RandNA $NumCompaniesWorked),col = 6 )
lines(density(Attrition_KNNRNA$NumCompaniesWorked),col = 4 )

```

# 4.3 Random Forest method for MNAR

```{r, echo=False}
## Miss Forest Imputation
forest_impNR <- missForest(Attrition_NoRandNA, xtrue = df, verbose = TRUE, variablewise = TRUE)
forest_impNR$ximp #valores imputados
forest_impNR$OOBerror #Estimated OOB imputation error
forest_impNR$error #True imputation error

#NRSM es el error cuadrático medio normalizado y se usa para representar el error derivado 
#de los valores continuos

#PFC (Proporcion de clasificados falsamente) se usa para representar el error derivado de 
#imputar las variables categoricas

##Se puede observar como el mayor error de prediccion corresponde a las variables Age y Hourlyrate
#sobre todo este ultimo
Conf2x2 = matrix(c(1:4), nrow=2, byrow=FALSE)
layout(Conf2x2)
layout.show(4)


plot(density(NONA$Age))
lines(density(forest_impNR$ximp$Age),col = 2 )
lines(density(Attrition_KNNNR$Age),col = 3 )
plot(density(NONA$HourlyRate))
lines(density(forest_impNR$ximp$HourlyRate),col = 2 )
plot(density(NONA$NumCompaniesWorked))
lines(density(forest_impNR$ximp$NumCompaniesWorked),col = 2 )
plot(density(NONA$YearsAtCompany))
lines(density(forest_impNR$ximp$YearsAtCompany),col = 2 )

```

# 4.4 Random Forest method for MCAR

```{r, echo=False}
forest_impRNA <- missForest(Attrition_missRand, xtrue = df, verbose = TRUE, variablewise = TRUE)
forest_impRNA$ximp #valores imputados
forest_impRNA$OOBerror #Estimated OOB imputation error
forest_impRNA$error #True imputation 
#PLOTS 
Conf2x2 = matrix(c(1:4), nrow=2, byrow=FALSE)
layout(Conf2x2)
layout.show(4)


plot(density(NONA$Age))
lines(density(forest_impRNA$ximp$Age),col = 2 )
lines(density(Attrition_KNNNR$Age),col = 3 )
plot(density(NONA$HourlyRate))
lines(density(forest_impRNA$ximp$HourlyRate),col = 2 )
plot(density(NONA$NumCompaniesWorked))
lines(density(forest_impRNA$ximp$NumCompaniesWorked),col = 2 )
plot(density(NONA$YearsAtCompany))
lines(density(forest_impRNA$ximp$YearsAtCompany),col = 2 )
```

# 4.5 mice method for MNAR

```{r , echo=False}
##MICE
Attrition_mice <- mice(Attrition_NoRandNA, m=5)
Attition_miceNR <- complete(Attrition_mice,5)

Conf2x2 = matrix(c(1:4), nrow=2, byrow=FALSE)
layout(Conf2x2)
layout.show(4)

plot(density(df$Age))
lines(density(Attition_miceNR$Age),col = 2 )
plot(density(df$HourlyRate))
lines(density(Attition_miceNR$HourlyRate),col = 2 )
plot(density(df$NumCompaniesWorked))
lines(density(Attition_miceNR$NumCompaniesWorked),col = 2 )
plot(density(df$YearsAtCompany))

```

# 4.6 mice method for MCAR

```{r , echo=False}
#MICE
Attrition_mice2 <- mice(Attrition_missRand, m=5)
Attition_miceMissR <- complete(Attrition_mice2,5)

Conf2x2 = matrix(c(1:4), nrow=2, byrow=FALSE)
layout(Conf2x2)
layout.show(4)

plot(density(df$Age))
lines(density(Attition_miceMissR$Age),col = 2 )
plot(density(df$HourlyRate))
lines(density(Attition_miceMissR$HourlyRate),col = 2 )
plot(density(df$NumCompaniesWorked))
lines(density(Attition_miceMissR$NumCompaniesWorked),col = 2 )
plot(density(df$YearsAtCompany))
## comparar los dos mice
plot(density(Attition_miceMissR$Age))
lines(density(df$Age), col = 3)
lines(density(Attition_miceNR$Age), col= 2)

```

# 5. Final Plots and Graphs

```{r, echo=False}
## FINAL PLOTS 
## NUMERICAL
#NO RANDOM
ggplot() + 
  geom_density(data = df, aes(x = NumCompaniesWorked), colour = "green")+
  geom_density(data = df_imp_knn5NR, aes(x = NumCompaniesWorked), colour = "yellow")+ 
  geom_density(data = forest_impNR$ximp,  aes(x = NumCompaniesWorked), colour = "red")+
  geom_density(data = Attition_miceNR,  aes(x = NumCompaniesWorked), colour = "blue")
# random
ggplot() + 
  geom_density(data = df, aes(x = Age), colour = "green")+
  geom_density(data = df_imp_knn5RandNA, aes(x = Age), colour =  "yellow")+ 
  geom_density(data = forest_impRNA$ximp,  aes(x = Age), colour = "red")+
  geom_density(data = Attition_miceMissR,  aes(x = Age), colour = "blue")
# COMPARATION OF DIFFERENT missing types
#KNN
ggplot() + 
  geom_density(data = df, aes(x = Age), colour = "green")+
  geom_density(data = df_imp_knn5RandNA, aes(x = Age), colour =  "yellow")+ 
  geom_density(data = df_imp_knn5NR,  aes(x = Age), colour = "red")
#RF
ggplot() + 
  geom_density(data = df, aes(x = Age), colour = "green")+
  geom_density(data = forest_impRNA$ximp, aes(x = Age), colour =  "yellow")+ 
  geom_density(data = forest_impNR$ximp,  aes(x = Age), colour = "red")
#MICE
ggplot() + 
  geom_density(data = df, aes(x = Age), colour = "green")+
  geom_density(data = Attition_miceMissR, aes(x = Age), colour =  "yellow")+ 
  geom_density(data = Attition_miceNR,  aes(x = Age), colour = "red")
## CATEGORICAL
# NO RANDOM 
col1 <- c(rep("Original",nrow(df)),rep("Knn",nrow(df_imp_knn5NR)),rep("RandomForest",nrow(forest_impNR$ximp)),rep("MICE",nrow(Attition_miceNR)))
col2 <- c(df$Gender,df_imp_knn5NR$Gender,forest_impNR$ximp$Gender,Attition_miceNR$Gender)

outcome_data <- cbind(col1, col2)

flip_results <- data.frame(Imputation = factor(col1, levels = c("Original", "Knn", "RandomForest", "MICE")),
                           Result = col2)
results <- data.frame(table(flip_results))

ggplot(data = results, aes(x = Result, y = Freq, fill = Imputation )) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
  geom_text(aes(label = Freq), fontface = "bold", vjust = 1.5,
            position = position_dodge(.9), size = 4) +
  labs(x = "\n Gender", y = "Frequency\n", title = "\n Gender Comparison \n") +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.title.x = element_text(face="bold", colour="red", size = 12),
        axis.title.y = element_text(face="bold", colour="red", size = 12),
        legend.title = element_text(face="bold", size = 10))


#RANDOM
col1 <- c(rep("Original",nrow(df)),rep("Knn",nrow(df_imp_knn5RandNA)),rep("RandomForest",nrow(forest_impRNA$ximp)),rep("MICE",nrow(Attition_miceMissR)))
col2 <- c(df$Gender,df_imp_knn5RandNA$Gender,forest_impRNA$ximp$Gender,Attition_miceMissR$Gender)

outcome_data <- cbind(col1, col2)

flip_results <- data.frame(Imputation = factor(col1, levels = c("Original", "Knn", "RandomForest", "MICE")),
                           Result = col2)
results <- data.frame(table(flip_results))

ggplot(data = results, aes(x = Result, y = Freq, fill = Imputation )) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)  +
  geom_text(aes(label = Freq), fontface = "bold", vjust = 1.5,
            position = position_dodge(.9), size = 4) +
  labs(x = "\n Gender", y = "Frequency\n", title = "\n Gender Comparison \n") +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.title.x = element_text(face="bold", colour="red", size = 12),
        axis.title.y = element_text(face="bold", colour="red", size = 12),
        legend.title = element_text(face="bold", size = 10))
```
