---
title: "Preprocessing"
author: ""
date: "2022-10-08"
output: html_document
---

TODO: EXPLANATION OF THIS PART

```{r, setup, include=FALSE, echo=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
requiredPackages <- c("ggplot2","rapportools","plyr","cowplot", "ggpubr", "corrplot", "corrplot", "mlbench", "caret", "FactoMineR", "ggcorrplot")
packages.check <- lapply(requiredPackages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

PROJECT_PATH="~/Documents/UPC/1Q/MultivariateAnalysis_MVVA/Labs/MultivariantAnalysis_IBMDataset"

setwd(PROJECT_PATH)

```

# Content Table

1.  Introduction

    -   1.1 Load Data

    -   1.2 DataFrame Basic Information

2.  Formatting issues

3.  Feature selection

    -   3.1 Categorical

        -   3.1.1 Contingency Tables

    -   3.2 Numerical

4.  Outliers Detection

    -   4.1 Uni-variant

    -   4.2 Multi-variant

5.  Derivation of New Variables

6.  Rank variables

7.  Save the new datasets

# 1. Introduction

## 1.1 Load Data and useful functions

```{r, echo=TRUE}
#df <- read.csv("C:/Users/jordi/Desktop/mva/data_mva.csv")
df <- read.csv("./Data/IBM_Employees.csv")

plotFreq <- function(df, feature, xlabel) {
  plot<-ggplot(df, aes(x=feature)) +
        geom_bar(aes(y =(..count..)/sum(..count..))) +
        geom_text(aes( label = scales::percent((..count..)/sum(..count..)),
                   y= ((..count..)/sum(..count..)) ), stat= "count", vjust = 1.25, color="#ffffff", size=3) +
        labs(x=xlabel, y='Percent')+
        scale_y_continuous(labels = scales::percent)
  return(plot)
}

plotPie <- function(feature, varName) {
  nb <- count(feature)[2];
  per <- c(round(nb/sum(nb)*100,2))$freq;
  names <- levels(feature);
  labs <- paste(names, per, "%");
  #jpeg(paste(plot_path, data_names[i], ".jpeg", sep = ""), height=800, width=1000)
  plot<-pie(summary(feature), labels = labs, main=paste("Pie Chart of ",varName));
  return(plot)
}

```

## 1.2 Dataframe basic information

Show a summary of the current dataframe.

```{r, echo=TRUE}
str(df)

```

# 2. Formatting issues

-   Check that Categorical variables are correctly processed by R, so all interpreted as factors.
-   Set short-meaningful level names for the categorical variables.
-   Reduce number of levels if has sense.

```{r, echo=False}

df$NumCompaniesWorked=df$NumCompaniesWorked+1

df$Attrition=as.factor(df$Attrition)

df$BusinessTravel=as.factor(df$BusinessTravel)

df$Department=as.factor(df$Department)

# Reduce number of levels from 29 to 7
df$DistanceFromHome=as.factor(df$DistanceFromHome)
plotFreq(df, df$DistanceFromHome, 'Distance From Home')
levels(df$DistanceFromHome) <- c("1", "1", "2", "2", "2", "2", "3", "3", "3", "3", "4", "4", "4", "4", "4", "5", "5", "5", "5", "5", "5", "5", "6", "6", "6", "6", "6", "7", "7")
plotFreq(df, df$DistanceFromHome, 'Distance From Home')

df$Education=as.factor(df$Education)
levels(df$Education) <- c('Below College', 'College', 'Bachelor', 'Master', 'Doctor')

df$EducationField=as.factor(df$EducationField)

df$EnvironmentSatisfaction=as.factor(df$EnvironmentSatisfaction)
levels(df$EnvironmentSatisfaction) <- c('Low', 'Medium', 'High', 'Very High')

df$Gender=as.factor(df$Gender)

df$JobInvolvement=as.factor(df$JobInvolvement)
levels(df$JobInvolvement) <- c('Low', 'Medium', 'High', 'Very High')

df$JobLevel=as.factor(df$JobLevel)
levels(df$JobLevel) <- c('Intern', 'Junior', 'Senior', 'Manager', 'Director')

# Reduce number of levels from 9 to 7
df$JobRole=as.factor(df$JobRole)
plotFreq(df, df$JobRole, 'Job Role')
levels(df$JobRole) <- c("Representatives", "H.R.", "Lab. Tech.", "Manager", "Directors", "Directors", "Research Sci.", "Sales Executive", "Representatives")
plotFreq(df, df$JobRole, 'Job Role')

df$JobSatisfaction=as.factor(df$JobSatisfaction)
levels(df$JobSatisfaction) <- c('Low', 'Medium', 'High', 'Very High')

df$MaritalStatus=as.factor(df$MaritalStatus)

df$Over18=as.factor(df$Over18)

df$OverTime=as.factor(df$OverTime)

df$PerformanceRating=factor(df$PerformanceRating, levels = c(1, 2, 3, 4))
levels(df$PerformanceRating) <- c('Low', 'Good', 'Excellent', 'Outstanding')

df$RelationshipSatisfaction=as.factor(df$RelationshipSatisfaction)
levels(df$RelationshipSatisfaction) <- c('Low', 'Medium', 'High', 'Very High')

df$StockOptionLevel=as.factor(df$StockOptionLevel)
levels(df$StockOptionLevel) <- c('No', 'Few', 'Medium', 'High')

df$WorkLifeBalance=as.factor(df$WorkLifeBalance)
levels(df$WorkLifeBalance) <- c('Bad', 'Good', 'Better', 'Best')

```

# 3. Feature selection

Evaluation of the relevant variables - Priorization and ranking under different criteria - Feature weighting - Elimination of irrelevant variables - Feature selection

## 3.1 Categorical variables

```{r, echo=False}

factorDf <- df[sapply(df, is.factor)]

# Binary Categorical Variables
plotPie(df$Attrition, 'Attrition')
plotPie(df$Gender, 'Gender')
plotPie(df$Over18, 'Over 18')
plotPie(df$OverTime, 'Over time')

# Categorical
p0<-plotFreq(df, df$BusinessTravel, 'Business travel')
p1<-plotFreq(df, df$Department, 'Department')
p2<-plotFreq(df, df$DistanceFromHome, 'Distance from home')
p3<-plotFreq(df, df$Education, 'Education')
plot_grid(p0, p1, p2, p3, nrow=2, ncol=2)

p0<-plotFreq(df, df$EducationField, 'Education field')
p1<-plotFreq(df, df$EnvironmentSatisfaction, 'Environment satisfaction')
p2<-plotFreq(df, df$JobInvolvement, 'Job involvement')
p3<-plotFreq(df, df$JobLevel, 'Job Level')
plot_grid(p0, p1, p2, p3, nrow=2, ncol=2)


plotFreq(df, df$JobRole, 'Job Role')

p0<-plotFreq(df, df$JobSatisfaction, 'Job Satisfaction')
p1<-plotFreq(df, df$MaritalStatus, 'Marital status')
p2<-plotFreq(df, df$PerformanceRating, 'Performance Rating')
p3<-plotFreq(df, df$RelationshipSatisfaction, 'Relationship Satisfaction')
plot_grid(p0, p1, p2, p3, nrow=2, ncol=2)

# The majority (+83%) of the employees have any or few stocks of the company, MAYBE RELATED WITH ATTRITION?
p0<-plotFreq(df, df$StockOptionLevel, 'Stock option level')
p1<-plotFreq(df, df$WorkLifeBalance, 'Work-life balance')
plot_grid(p0, p1, nrow=1, ncol=2)


df <- subset(df, select = -c(Over18))

factorDf <- df[sapply(df, is.factor)]


```

### 3.1.1 Contingency table

```{r, echo=False, warning=FALSE}

# for all combination between two categorical variable, we make a chisq.test (independance test for categorical var) and we keep variable than are dependant (p-value of the chisq test < 0.05 => we are sure than there are dependant with 95%)

i=1
for (var1 in factorDf[-length(factorDf)]) {
  factorDf3 <- factorDf[-c(1:i)]
  j=1
  for(var2 in factorDf3) {
    if(attributes(factorDf[i])$names != "Attrition" && attributes(factorDf3[j])$names != "Attrition"){
      chisq_test <- chisq.test(var1, var2)
      if(chisq_test$p.value < 0.01) {
        print(paste0(attributes(factorDf[i])$names, " | ", attributes(factorDf3[j])$names, " | p-value ", chisq_test$p.value))
        #print(table(var1,var2))
      }
    }
    j=j+1
  }
  i=i+1
}

ggplot(df, aes(Department, ..count..)) + geom_bar(aes(fill = EducationField), position = "dodge")
ggplot(df, aes(Department, ..count..)) + geom_bar(aes(fill = JobLevel), position = "dodge")
ggplot(df, aes(Department, ..count..)) + geom_bar(aes(fill = JobRole), position = "dodge")
ggplot(df, aes(Education, ..count..)) + geom_bar(aes(fill = JobLevel), position = "dodge")
ggplot(df, aes(EducationField, ..count..)) + geom_bar(aes(fill = JobLevel), position = "dodge")
ggplot(df, aes(EducationField, ..count..)) + geom_bar(aes(fill = JobRole), position = "dodge")
ggplot(df, aes(JobLevel, ..count..)) + geom_bar(aes(fill = JobRole), position = "dodge")
ggplot(df, aes(JobLevel, ..count..)) + geom_bar(aes(fill = StockOptionLevel), position = "dodge")
ggplot(df, aes(MaritalStatus, ..count..)) + geom_bar(aes(fill = StockOptionLevel), position = "dodge")

```

## 3.2 Numerical Variables

```{r, echo=False}

nonFactorDf <- df[!sapply(df, is.factor)]

#EmployeeNumber
ggplot(df, aes(x = 1:nrow(df), y = EmployeeNumber)) +
  geom_point()+
  labs(x = "Table Index")

sd(df$EmployeeNumber)

# Como se puede observar con la grafica de los valores respecto a su indice en la tabla,
# estamos frente a un indice para cada empleado sin un significado m치s alla del numerico y
# por lo tanto debemos eliminar la columna.


#EmployeeCount
ggplot(df, aes(x = 1:nrow(df), y = EmployeeCount)) +
  geom_point()+
  labs(x = "Table Index")

sd(df$EmployeeCount)

#As we can see in the plot of the variable and the sd, all the values of the table correspond to 1.


#RateVariables 

pairs(df[,c("DailyRate", "HourlyRate", "MonthlyRate")])
pairs(df[,c("DailyRate", "HourlyRate", "MonthlyRate", "YearsAtCompany","MonthlyIncome")])
# podemos ver que la relacion entre ellas es nula

ggplot(df, aes(x = MonthlyRate, y = MonthlyIncome)) +
  geom_point()

#df$DailyRate / df$HourlyRate

ggplot(df, aes(x = 1:nrow(df), y = DailyRate / HourlyRate)) +
  geom_point()+
  labs(x = "Table Index")



corr <- round(cor(nonFactorDf), 1)


#ggcorrplot(corr, hc.order = TRUE, type = "lower",
#   lab = FALSE)

#ggcorrplot(corr, method = "circle")



#Remove redundantcy with lineal correlation study

correlationMatrix <- cor(nonFactorDf)

print(correlationMatrix)
# find attributes that are highly corrected (ideally >0.75)
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.75)
# print indexes of highly correlated attributes
print(highlyCorrelated)

#Highly corralated variables:
# - Total Working years - monthly income
# - Years with current manager - Years at the company - Years current role
# We should think of eliminating the variables that are highly correlated.
# Probably doing and stadistic test would be a good idea to know if we have to
# delete them or no.

df <- subset(df, select = -c(EmployeeNumber, EmployeeCount, StandardHours))

nonFactorDf <- df[!sapply(df, is.factor)]



# Analysis of TrainingTimesLastYear
hist(df, df$TrainingTimesLastYear, 1, "TrainingTimesLastYear")
var(df$TrainingTimesLastYear)

count(df$TrainingTimesLastYear)


# Analysis of PercentSalaryHike

hist(df, df$PercentSalaryHike, 1, "PercentSalaryHike")


count(df$PerformanceRating)

# strongly correlated
ggplot(df) + geom_boxplot(aes(x = PercentSalaryHike, y = PerformanceRating)) 




ggplot(df,aes(x=TotalWorkingYears, y=MonthlyIncome)) +
  geom_point() +
  geom_smooth(method='lm') 



ggplot(df,aes(x=YearsWithCurrManager , y=YearsAtCompany)) +
  geom_point() +
  geom_smooth(method='lm') 


ggplot(df,aes(x=YearsInCurrentRole, y=YearsWithCurrManager)) +
  geom_point() +
  geom_smooth(method='lm') 




hist(df, df$MonthlyIncome, 350, "MonthlyIncome")

```

# 4. Outliers Detection

## 4.1 Uni-Variant

```{r, echo=False}

#Boxplots
box_plot <- function(data, column, name)
{
  return(ggplot(data, aes(column)) + 
           geom_boxplot()+
           labs(x = name))
}

bp_PercentSalaryHike <- box_plot(df, df$PercentSalaryHike, "PercentSalaryHike")
bp_TotalWorkingYears <- box_plot(df, df$TotalWorkingYears, "TotalWorkingYears")
bp_TrainingTimesLastYear <- box_plot(df, df$TrainingTimesLastYear, "TrainingTimesLastYear")
bp_YearsAtCompany <- box_plot(df, df$YearsAtCompany, "YearsAtCompany")
bp_YearsInCurrentRole <- box_plot(df, df$YearsInCurrentRole, "YearsInCurrentRole")
bp_YearsSinceLastPromotion <- box_plot(df, df$YearsSinceLastPromotion, "YearsSinceLastPromotion")
bp_YearsWithCurrManager <- box_plot(df, df$YearsWithCurrManager, "YearsWithCurrManager")
bp_MonthlyIncome <- box_plot(df, df$MonthlyIncome, "MonthlyIncome")

figure_bp <- ggarrange(bp_PercentSalaryHike, bp_TotalWorkingYears, bp_TrainingTimesLastYear,
                    bp_YearsAtCompany, bp_YearsInCurrentRole,
                    bp_YearsSinceLastPromotion, bp_YearsWithCurrManager, 
                    bp_MonthlyIncome,
                    ncol = 2, nrow = 4)
figure_bp



#Histograms
hist <- function(data, column, bw, name)
{
  return(ggplot(data, aes(column)) + 
           geom_histogram(aes(y = ..density..), colour="black", fill="white", binwidth = bw)+
           geom_density(lwd = 1, colour = 4, fill = 4, alpha = 0.2)+
           labs(x = name))
}

hist_PercentSalaryHike <- hist(df, df$PercentSalaryHike, 1, "PercentSalaryHike")
hist_TotalWorkingYears <- hist(df, df$TotalWorkingYears, 1, "TotalWorkingYears")
hist_TrainingTimesLastYear <- hist(df, df$TrainingTimesLastYear, 1, "TrainingTimesLastYear")
hist_YearsAtCompany <- hist(df, df$YearsAtCompany, 1, "YearsAtCompany")
hist_YearsInCurrentRole <- hist(df, df$YearsInCurrentRole, 1, "YearsInCurrentRole")
hist_YearsSinceLastPromotion <- hist(df, df$YearsSinceLastPromotion, 1, "YearsSinceLastPromotion")
hist_YearsWithCurrManager <- hist(df, df$YearsWithCurrManager, 1, "YearsWithCurrManager")
hist_Age <- hist(df, df$Age, 1, "Age")
hist_DailyRate <- hist(df, df$DailyRate, 150, "DailyRate")
hist_HourlyRate <- hist(df, df$HourlyRate, 50, "HourlyRate")
hist_MonthlyIncome <- hist(df, df$MonthlyIncome, 700, "MonthlyIncome")
hist_MonthlyRate <- hist(df, df$MonthlyRate, 650, "MonthlyRate")
hist_NumCompaniesWorked <- hist(df, df$NumCompaniesWorked, 1, "NumCompaniesWorked")

#juntar plots en ggplot2
figure_hist_1 <- ggarrange(hist_PercentSalaryHike, hist_TotalWorkingYears, hist_TrainingTimesLastYear,
                      hist_YearsAtCompany, hist_YearsInCurrentRole, hist_YearsSinceLastPromotion, 
                      ncol = 3, nrow = 2)
figure_hist_1

figure_hist_2 <- ggarrange(hist_YearsWithCurrManager, hist_MonthlyIncome,
                      hist_Age, hist_DailyRate, hist_MonthlyRate, 
                      hist_NumCompaniesWorked,
                      ncol = 3, nrow = 2)
figure_hist_2

```

## 4.2 Multi-variant

```{r, echo=False}

mdi = mahalanobis(nonFactorDf,center=apply(nonFactorDf,2,mean),cov=var(nonFactorDf))
plot(density(mdi), main="Density curve of Mahalanobis distance")
cutoff <- qchisq(p = 0.99 , ncol(nonFactorDf))

# Create new DataFrame and set a new categorical variable to identify which are multivariate outliers or not

tmpDF<-df
tmpDF$MultivariateOutlier<-0
tmpDF[mdi>cutoff,]$MultivariateOutlier<-1
tmpDF$MultivariateOutlier<-as.factor(tmpDF$MultivariateOutlier)

monthlyIncomeMean<-mean(df$MonthlyIncome)

## Display mean for Whole dataset, non multivariate outliers and multivariate outliers
apply(nonFactorDf, 2, mean)
apply(nonFactorDf[mdi<=cutoff,], 2, mean)
apply(nonFactorDf[mdi>cutoff,], 2, mean)

# Non Outliers individuals
dfMahalanobisNonOutliers<-df[mdi<=cutoff,]

# Outliers individuals
dfMahalanobisOutliers<-df[mdi>cutoff,]

# Subgroup found in Outliers individuals
dfMahalanobisOutliers2<-dfMahalanobisOutliers[dfMahalanobisOutliers$MonthlyIncome<monthlyIncomeMean,]
dfMahalanobisOutliers2[,c("Age","MonthlyIncome","NumCompaniesWorked","TotalWorkingYears","YearsAtCompany","YearsInCurrentRole","YearsSinceLastPromotion","YearsWithCurrManager")];


## Color By Outliers (Whole Dataset)
ggplot(tmpDF, aes(MonthlyIncome, Age, color=MultivariateOutlier))+
  geom_point()
ggplot(tmpDF, aes(MonthlyIncome, TotalWorkingYears, color=MultivariateOutlier))+
  geom_point()
ggplot(tmpDF, aes(Age, TotalWorkingYears, color=MultivariateOutlier))+
  geom_point()
ggplot(tmpDF, aes(PercentSalaryHike, YearsAtCompany, color=MultivariateOutlier))+
  geom_point()
ggplot(tmpDF, aes(JobRole, YearsAtCompany, color=MultivariateOutlier))+
  geom_point()
ggplot(tmpDF, aes(PercentSalaryHike, TotalWorkingYears, color=MultivariateOutlier))+
  geom_point()
ggplot(tmpDF, aes(JobRole, TotalWorkingYears, color=MultivariateOutlier))+
  geom_point()

## Color By JobRole (Only Outliers)
ggplot(dfMahalanobisOutliers, aes(MonthlyIncome, Age, color=JobRole))+
  geom_point()
ggplot(dfMahalanobisOutliers, aes(MonthlyIncome, TotalWorkingYears, color=JobRole))+
  geom_point()

```

# 5. Derivation of new variables

```{r, echo=False}


# Porcentaje de salario que le han agumentado cada a침o.
df$SalaryHikePerYear <- df$YearsAtCompany

df$SalaryHikePerYear[df$SalaryHikePerYear == 0] <- 1

df$SalaryHikePerYear <- df$PercentSalaryHike / df$SalaryHikePerYear


#A침os de media en saltar de empresa a empresa

df$TotalWorkingYears_aux <- df$TotalWorkingYears
df$TotalWorkingYears_aux[df$TotalWorkingYears_aux == 0] <- 1 

df$AverageYearsCompany <- df$TotalWorkingYears_aux / df$NumCompaniesWorked

df$TotalWorkingYears_aux <- NULL


#A침os para jubilarse
JubliationAge <- 66

df$AgeToJubilation <- JubliationAge - df$Age


# The employee has different managers in the company
df$DifferentManagers<-df$YearsAtCompany-df$YearsWithCurrManager
df$DifferentManagers[df$DifferentManagers>0]<-1
df$DifferentManagers<-as.factor(df$DifferentManagers)
levels(df$DifferentManagers)<-c("No","Yes")

ggplot(dfMahalanobisOutliers, aes(MonthlyIncome, TotalWorkingYears, color=JobRole))+
  geom_point()
```

# 6. Rank variables

```{r, echo=False}

factorDf <- df[sapply(df, is.factor)]
nonFactorDf <- df[!sapply(df, is.factor)]

tmpDf<-data.frame(nonFactorDf,factorDf['Attrition'])

control <- trainControl(method="repeatedcv", number=10, repeats=3)

model <- train(Attrition~., data=tmpDf, method="lvq", preProcess="scale", trControl=control)

importance <- varImp(model, scale=FALSE)

print(importance)

plot(importance)

```

# 7. Save the new datasets

```{r, echo=False}

# Preprocesed Dataset
save(df, file = "./Data/df_preprocessed_IBM_Employees.Rdata")

#################################
# PCA - MCA - MFA - ML Datasets #
#################################

# Dataset For PCA
df_PCA<-subset(df, select = -c(Department,EducationField))
df_PCA<-df_PCA[!sapply(df_PCA, is.factor)]
save(df_PCA, file = "./Data/df_preprocessed_PCA_IBM_Employees.Rdata")

# Dataset For MCA
df_MCA<-subset(df, select = -c(Department,EducationField))
df_MCA<-df_MCA[!sapply(df_MCA, is.factor)]
save(df_MCA, file = "./Data/df_preprocessed_MCA_IBM_Employees.Rdata")

# Dataset For MFA
#write.table(dd, file = "./Data/preprocessed_MFA_IBM_Employees.csv", sep = ",", na = "NA", dec = ".", row.names = FALSE, col.names = TRUE)

# Dataset For Machine Learning Models
#write.table(dd, file = "./Data/preprocessed_ML_Employees.csv", sep = ",", na = "NA", dec = ".", row.names = FALSE, col.names = TRUE)

```
